---
title: "Project5"
output: html_document

---




#Introduction: 
The Dataset that I use was about the quality of red wine and how it affect by
the elements that are in it. The dataset come from Cortez et al., who were 
trying to find the preference of wine . What Interest me about the wine data how much elements affect the overall quality of the wine. My plan is to find out 
more detail how the element affect the quality.


#Problem : 
I want to see how the element of the wine affect the quality of the 
wine.

```{r, message=FALSE, warning=FALSE, include=FALSE}

library(dplyr)
library(reshape2)
library(data.table)
library(scales)
library(ggplot2)
```

#Univariate Plot


The code Below insert the file wineQualityReds and place the infomation into 
redwine.Also get some basic infomation about the Dimension , Structure and basic
summary of the csv file. This give me a idea how to read the diagram and give me
a range to look at for each column.
```{r, include=FALSE}
redwine<-read.csv("wineQualityReds.csv")
```

```{r, echo=FALSE}
redwine <- subset(redwine, select = -X)
+dim(redwine)
```


```{r, echo=FALSE}


 str(redwine)
```


```{r, echo=FALSE}
summary(redwine)
```





The Histograms below show a representation of each column from the file:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = fixed.acidity))+ 
  geom_histogram()+scale_x_continuous()
```

  The diagram above is a representation of the fixed.acidity column. Something 
that interest me about he graph that it start around 4.5 and not 0. Second thing
i notice that there no fixed.acidity around 15.5.






```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = volatile.acidity ))+ 
  geom_histogram()
```
  
  The diagram above is representation of the volatile.acidity. The interesting 
about their a big jumb around .3 volatile and  a missing volatile.acidity around
1.3 and 1.5 volatile.acidity.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = citric.acid))+ 
  geom_histogram()
```
 
  The histogram above show the representation for citric.acid something 
interesting that their is a drop aroun .16 citric.acid but a big jump at .18
citric.acid.Also a their no citric.acid between .8 and .97.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = residual.sugar ))+ 
  geom_histogram()
```
  
  The graph above represent the residual.sugar column. The interesting that it 
positive scew but after the residual.sugar reach 3 it don't increase and then it 
reach 10 the tail start cutting.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = chlorides))+ 
  geom_histogram()
```
  
  The graph above represent chloride from the csv fie what interest me about 
thisgraph that their great drop close to .1 and the chloride don't start at 0 
but at .5.



```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = free.sulfur.dioxide ))+ 
  geom_histogram()
```
 
  The graph above is the a representation of free.sulfur.dioxide something i 
notice is that their alot of point that their a drop and then it stay steady 
then it increase a good example is around 18 to 28.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = total.sulfur.dioxide ))+ 
  geom_histogram()
```
  
   The graph above it a representation of  Total.Sulfur.Dioxide.The interesting
   part of this graph it would been positive scew but it increase at 80 to 90 
   total.sulfur.dioxide. We also see small  amout of total.sulfur.dioxide near
   300 we can belive does are outliers .
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = density ))+ 
  geom_histogram()
```

The graph above show a representation of density it a good representation of the 
a Histogram.Something that i notice that the tail start at .990 and it make 
sence since histogram don't touch zero.




```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = pH))+ 
  geom_histogram()
```

The graph above show a good representation of a Histogram for pH. Something I
notice that near 2.6 ,3.8.,and 3.9 the graph cut off and then reappers. 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = sulphates))+ 
  geom_histogram()
```

The histograph above it a representation of sulphates i notice that the graph 
start closer to 0 and doesn't go pass 2 in suplate but reaches 355 in the count 
around .6 sulphates 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = alcohol ))+ 
  geom_histogram()
```

The graph from about it a representation of alcohol something i notice that it 
reach it peak around 9.2 and slowly start dropping. Another thing i notice their
a outlier near 16 alcohol.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwine , aes(x = quality ))+ 
  geom_bar()
```

The Graph above is representation of the quality of the wine. Something that 
suprise me that their not 0,1,9, and 10 quality of wine and most wine quality 
are around 5 or 6. 


################################################################################

We are going to be looking into the Total sulfure dioxide and the level of
concetration. If the level is greater than 50ppm it can be
detect by nose and taste.In The line below you can see basic infomation about
the So2.
```{r, echo=FALSE}
summary(redwine$free.sulfur.dioxide)
```

In the line below I see that 546 wines are greater than 50ppm so they have 
SO2 and 1053 wines don't have SO2.
```{r, echo=FALSE}
dim(subset(redwine, total.sulfur.dioxide >50 ))
```

```{r, echo=FALSE}
dim(subset(redwine, total.sulfur.dioxide <= 50 ))
```


```{r, eval=FALSE, include=FALSE}


library(ggplot2)


#ggplot(data = redwine , aes(x = total.sulfur.dioxide ))+ 
#  geom_histogram()
```



```{r, echo=FALSE}
ggplot(data = redwine , aes (x = total.sulfur.dioxide))+
  geom_histogram(binwidth = 10)+scale_x_continuous()+
  xlab("Total Sulfur.Dioxide(ppm)")
```

################################################################################
After looking Total.Sulur.Dioxide i start wondering how that affect the overall
quality of the wine.So i decide to facet_wrap that to my code. Which you can see
below.

                                                  
```{r, echo=FALSE}
ggplot(data = redwine , aes (x = total.sulfur.dioxide))+
  geom_histogram(binwidth = 20)+scale_x_continuous()+ 
  facet_wrap(~quality,ncol = 2)
```


From the data above i notice that most Total.SUlfur.Dioxide have a quality of 
5 or a 6.

The information below show Min,1st Qu,Median,Mean, 3rd Qu, and Max for each 
quality type
```{r, echo=FALSE}

by(redwine$total.sulfur.dioxide, redwine$quality,summary)
```
```{r, echo=FALSE}
ggplot(data = redwine , aes (x = total.sulfur.dioxide))+
  geom_histogram(binwidth = 5)+scale_x_continuous()+ 
  facet_wrap(~quality,scales= 'free')
```

The graph above is facet_wrap with a free scale giving me a better idea of 
toal.sulfur.dioxide for each quality.Quality 5 and 6 a range from close to 0 
to 150. While Quality 3, 4, and 8 have more a spread out Total.Sulfur.Dioxide

```{r, echo=FALSE}
ggplot(data= redwine, aes (x = as.factor(quality)  ,y= total.sulfur.dioxide))+
  geom_boxplot()+xlab("Quality")+ylab("Total Sulfur Dioxide")
```

#Univeriate Analysis:
###Structure of the Dataset:
Variables :11 Obj:1599

Fixed.acidity ,Volatile.acidity,citric.acid.,residual.sugar,chlorides,
free.sulfur.dioxide,total.sulfur.dioxide.,density,pH,sulphates,alcohol,

Quality : Ranking 0 -10 (Worst to best)

Observation:
The mean of th quality is around 5.6.

Most wine have a quality of a 5 or a 6.

Most wine use less residual.sugar(less than 2.5)

Most have less than 100ppm (fix.acidty)the quality of wine is a 6


###What is the most interesting from the dataset?
Something that i notice in most of the data set of the wine have use less 
element what i mean that elements like fixed.acidity , volatile.acidity,
residual.sugar,chlorides,freesulfur.dioxide,toal.sulfur.dioxide,sulphates,and
alcohol start with small quantity of the element but have a huge count and more
the quantity of the element the less is the count. We can determin that each 
element affect the quality of the wine by have different reaction with each 
other if not having specific amount. 

###What is something strange i see in the database?
One strange thing i saw was citric.acid because when have close to 0 citric.acid
it have high level count but then it slowly start to drop then near .18 
citic.acid jump in count and then slowly start to decrease. I believe their 
must be a element that citric acid reaction we if the quantity is measure 
correctly. I use the correlation diagram to see which element have a strong
relationship.

###What can help specify my theory?
Well i look at the total.sulfur.dioxide and quality above and see that their is 
some kind relation with elements and quality if you look at the scale most
total.sulfur.dioxide have less than 100 ppm very few have anything getter than
than amount. I want to look at the two element and how they affect each other
relationship that have a positive correlation and a negative correlation. 


```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#ggplot(data = redwine, aes(x=quality))+
#  geom_freqpoly()+scale_x_continuous()+xlab("Quality")+
#  ylab("Total Sulfur Dioxide")
```



#Correlation
After looking into the relationship between quality and Total.Sulfur.Dioxide
I start wondering about what other elements have a correlation with each other.I 
created two corrplot to give me a better idea of correlation between the 
element. The reason i create two plot was that first plot with square gave me a 
good idea how each correleation between each element but it still didn't tell me
how close to -1 or 1 so i plot another so i can get those results.


```{r, echo=FALSE}
library(corrplot)
rd <- cor(redwine )

corrplot(rd, method = 'square' )
corrplot(rd , method = 'number')


```

#Bivariate: 
###Fix.acidity and Citric.Acid

After looking at the corr.plot from above i notice that fixed.acidity which is 
the acid involved in wine and citric.acid which is add freshness and flavour 
have a correlation of .67.I want to see more about the relationship so i decide
to create a scatter plot

```{r, echo=FALSE}
ggplot(aes(x = fixed.acidity, y = citric.acid),data= redwine)+
  geom_point()
```

The first think i know notice that their alot of overplotting between 
fixed.acidiy = 6 and 10. I'm going to use aplha to clean up the overplotting.


```{r, echo=FALSE}
ggplot(aes(x = fixed.acidity, y = citric.acid),data= redwine)+
  geom_point(alpha= 1/10)+scale_x_continuous()
```

I Clean the overploting the first thing i notice is that most of the citric.acid
and fixed.acidity is close to when fixed.acidity is close to 7 and citric.acid
is  at 0,

```{r, echo=FALSE}


cor.test(redwine$citric.acid, redwine$fixed.acidity)

```
The information from above is the correlation between citric.acid and 
fixed.acidity using the Peason method.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = fixed.acidity, y = citric.acid),data= redwine)+
  geom_point(alpha = 1/10)+xlim(4.5,quantile(redwine$fixed.acidity,.99))+
  geom_smooth(method = 'lm', color = 'red')
```




```{r, echo=FALSE}
quantile(redwine$fixed.acidity,.99)
```
In the graph above i want to see the 99 percent tile so i can have a better 
idea of the graph without the outliers. The last fixed acidity that is in the 
99% is 13.3 anything beyond that is most likely a outlier.
```{r, echo=FALSE}

fixed_group <- group_by(redwine,fixed.acidity)
fixedsum    <- summarise(fixed_group,
                         mean_acid =mean(citric.acid),
                         median_acid = median(citric.acid),
                         max_acid = max(citric.acid),
                         min_acid = min(citric.acid),
                         n = n())
fixedsum <- arrange(fixedsum)
head(fixedsum)
```
The table above is a group_by fixed.acity and citric.acid.I wanted to get the 
mean ,median,max, min and total of the citric acid when fixed.acidity is at a 
a specific point.





###Fixed.Acidity and Density

While working with Fixed.acidity I notice that Density have a correlation of .68
and pH have correlation of -.67 with Fixed.acidity and I wanted to see more 
about their relatioship.Density is the density of water and pH is how acidic the
wine is. 

```{r, echo=FALSE}
ggplot(aes(x =fixed.acidity, y = density), data = redwine)+geom_point()
```

In the graph above I'm plotting between fixed.acidity and density. The problem
with this is that there is overploating when fixed.acidity reaches 8. I'm going 
to need to clean that before going any futher.


```{r, echo=FALSE}
ggplot(aes(x =fixed.acidity, y = density), data = redwine)+
  geom_point(alpha = 1/10)
```

I clean the graph from above you know that most of the plot are around 
fixed.acidity is around 8 and the density is around.9975.Now I'm wondering how
the quality is affect.
```{r, echo=FALSE}
cor.test(redwine$density, redwine$fixed.acidity)
```




```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = fixed.acidity, y = density),data= redwine)+
  geom_point(alpha = 1/10)+xlim(4.5,quantile(redwine$fixed.acidity,.99))+
  geom_smooth(method = 'lm', color = 'red')

```



###Fixed.Acidity and pH 

After doing the correlation for  citric.acid and density seeing that have a 
positive I want to compare it to element that have a negative correlation.I want
to look at pH and fixed.acidity since i know that they have a negative 
correlation.pH is the acid the wine is.

```{r, echo=FALSE}
ggplot(aes(x = fixed.acidity, y = pH),data= redwine)+
  geom_point()
```

I create a scatter plot the first thing i notice is their alot overplotting 
around 7 and 8. It better to clean the data so we can have a better 
understanding of the graph. 


```{r, echo=FALSE}
ggplot(aes(x = fixed.acidity, y = pH),data= redwine)+
  geom_point(alpha = 1/20)
```

Now that he graph is clean you notice that most of the point are around 7 for 
fixed.acidity it seem that he relationship stay mostly in that area. 
```{r, echo=FALSE}
cor.test(redwine$pH, redwine$fixed.acidity)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = fixed.acidity, y = pH),data= redwine)+
  geom_point(alpha = 1/10)+xlim(0,quantile(redwine$fixed.acidity,.99))+
  geom_smooth(method = 'lm', color = 'red')
```

#Bivariate Analysis:
###What are the interesting relationship did you see ?
Their alot of overplotting closer to the center of the graph (Fixed.Acidity/
Density and Fixed.Acidity/pH) .Fix.acidity and citric.acid is different because
most of the overplotting was found when citric acid is 0. This could mean that
fixed.acidity could have a strong relationship with citric.acid when it at 0.

###Strongest Relationship found? 
According to correlation diagram fixed.acidity have a strong relation with both
citric.acid and denstiy at .67. When i look at fixed.acidity and density scatter
plot graph it look like it have strong reaching since more of the scatter plot 
overplotting each other but still have a few that are spread out.Also 
fixed.acidity have a negative relation with pH when you look at the graph you 
can tell since it going in a negative linear line.



#Multivariate:
###Fixed.Acidity and Citric.acid with quality 
```{r, echo=FALSE}
ggplot(aes(x = fixed.acidity, 
           y = citric.acid , color = as.factor(quality)), 
       data = redwine)+geom_point()



ggplot(aes(x = fixed.acidity, 
           y = citric.acid  , color = factor(quality)), 
       data = redwine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality'))+theme_dark()
```

In the graph above i wanted to see the quality of the citric.acid and
fixed.acidity the lighter the color the worst is the quality , the darker the 
color the better the quality.SOmething i notice from this graph is that most
of the 3 and 4 quality are below .25 ctiric acid and most 7 and 8 quality are 
above .25 citric with 8 fixed.acidity

The information below we can see how many of each type of quality is there 
between 0 - 10.
```{r, echo=FALSE}

quality <- redwine$quality
summary(quality)
table(quality)

redwine$quality_bucket <- cut(redwine$quality, breaks = c(2,4,5,6,8,9))

```

```{r, eval=FALSE, include=FALSE}
#ggplot(aes(x = fixed.acidity, y = citric.acid),data= redwine)+
#  geom_point(aes(color = quality_bucket))+geom_smooth(method = 'lm' , 
#  color = 'red')
 
```
 
 In the infromation above i break down the quality to see how many are in each
 quality type.  0,1,2,9,10 are not show because their no value in the data. 

##Fixed.Acidity and Density with quality 

```{r, echo=FALSE}

ggplot(aes(x = fixed.acidity, 
           y = density  , color = as.factor(quality)), 
       data = redwine)+geom_point()



ggplot(aes(x = fixed.acidity, 
           y = density  , color = factor(quality)), 
       data = redwine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality'))+theme_dark()
```

The graph above now show me the quality of the density and fixed.acity.Something
I notice from the graph above is that closer you get to .995 from 1 in density
the better quality get and most of them are between 6 and 8 fixed,acidity



```{r, eval=FALSE, include=FALSE}
#ggplot(aes(x =fixed.acidity, y = density), data = redwine)+
#geom_point(aes(color = quality_bucket))+geom_smooth(method = 'lm', 
#color = 'red')
```




##Fixed.Acidity and pH with quality 

```{r, echo=FALSE}

ggplot(aes(x = fixed.acidity, 
           y = pH  , color = as.factor(quality)), 
       data = redwine)+geom_point()


ggplot(aes(x = fixed.acidity, 
           y = pH  , color = factor(quality)), 
       data = redwine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality'))+theme_dark()
```

I place quality into graph to see how pH and fixed.acidity affect the quality of 
the wine sinne something i notice that i going in a negative direction and 
quality 4,5,6,7,8 all intercept at 7 fixed.acidity and when pH reach around 3.8.
This interesting because it mean that another element that not pH and fixed 
affect the quality of the wine.


#Multivariate Analysis:
###What are the interesting relationship did you see ?
Something that I notice between all three graph is that one point in each graph
they have a intercept or come close to intercept with each other.Two of the 
graph have positive linear slope but that can be due to a postive correlation 
that they have with each other. Something i was suprise to see that all three
have overplotting  instead a spread out relationship.

###Strongest Relationship found?
It between  fixed.acidity with citric.acid or with Density since their 
correlation were the same. Since we add quality to the mix it just let us know 
when does the quality of the wine are plotted between fixed.acidity with
citric.acid or density.


```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#ggplot(aes(x = fixed.acidity, y = pH),data= redwine)+
#  geom_point(aes(color = quality_bucket))+geom_smooth(method = 'lm' , 
#  color = 'red')
```






#Three Interesting Graph
Will working on the project I found these Three graph to be the most interesting

##Graph 1:
```{r, echo=FALSE}
ggplot(data = redwine , aes (x = total.sulfur.dioxide ))+
  geom_histogram(binwidth = 5)+scale_x_continuous()+ 
  facet_wrap(~quality,scales= 'free')+
  labs(title="Total Sulfur Dioxide fracet_wrap with Quality")+
  xlab("Total Sulfur Dioxide units: mg /dm^3")
```

The reason i beleive this graph is import is because it show me it break down
how total.sulfur.dioxide with the quality of the wine. When the quality is 3 
most of the wine is less than 50ppm meaning now there no S02 in the wine. When 
you get quality of the wine is 5 you that reach it peak near 25 but the tail
slowly start dropping this can mean that SO2 have a higher detection.Lastly I 
notice that their are no wine rating of 0,1,2,9,10.

##Graph 2: 
```{r, echo=FALSE}
ggplot(aes(x = fixed.acidity, 
           y = density  , color = factor(quality)), 
       data = redwine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality'))+
  xlab("Fixed Acidity  Units: g /dm^3")+ylab("Density Units:g/m^3")+theme_dark()
```

In the graph above is important because it show how the density and Fixed 
Acidity afffect the quality of the wine. Somthing I notice is when the Fixed 
Acidity is between 13g/dm^3 and 14g/dm^3 the quality of the wine start to get to
each other. If the sample was larger we could determine if it ever intercept 
with each other. Another is the most the better quality wine is when the density 
is .992g/m^3 and the worst is closer to .997 g/m^3.


##Graph 3

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = fixed.acidity, y = citric.acid),data= redwine)+
  geom_point(alpha = 1/10)+xlim(4.5,quantile(redwine$fixed.acidity,.99))+
  geom_smooth(method = 'lm', color = 'red')+
  labs(title = "How Citric.acid and Fixed.acidy affect Quality")+
  ylab("Citric.Acid  Units :g /dm^3 ")+xlab("Fixed Acidity Units: g/dm^3")

```

The last graph is very import because it Bivariate between citric.acid and
fixed.acidity. The reason this is important because it show the 99% of the graph
removing the outliers. I use the alpha function to remove some of the 
overplotting this show me that most of fixed.acidity is around 7.5 with a
citric.acid of .25. We also that see that their a linear line that positve
giving it a positive correlation.


#Limitation:

A couple of limitation for was understanding the each element and seeing how
they affect the wine or each element. Luckly i saw that the file come with a
explanation of each element but don't really give me a idea how each element 
with each other. So i start with Total.Sulfur.Dioxide i know that it can affect
the quality of the wine if the ppm is to high. I wanted to have a better idea 
how each element relate to each other i decide to create 2 correlation graph 
to have abetter idea.

#Ideas:
I had a couple ideas while working with these set of data. The first was data
could contain infomation how well each wine sold and then compare it to the 
quality of the wine and each element. The problem with this method is that maybe
some wine are the same but sold at a different cost at different stores. Another
is to see how Temperature affect the element and the overal quality of the wine.
The problem is that i see that all the wine must have the same amount of time
in that temperature it would be hard to tell because they might have diferent 
makers.

#Conclusion:
This project help me understand the use of R and graph for data science.Graph 
are really important to this field because it help the reader have a better 
picture when you explain the research you have done. 














